Writing Trash Bin

### 2024-05-01 paper.qmd
small edit + figures for background

# Background

From entertainment choices to news articles to insurance plans, the modern citizen is so over run with information in every aspect of their life it can be overwhelming. In this overflow of information, tools that can effectively summarise information down into simple and clear ideas become more valuable. Information visualisations remain one of the most powerful tools for fast and reliable science communication.

There are many stages in our analysis that might benefit from the power of data visualisation, however this does not mean it is appropriately utilized or done with with effectiveness. (@Vanderplas2020 has helpful advice here.) Visualization is an important step in exploratory data analysis and it is often utilised to **learn** what is important about a data set. The importance of data-driven discovery has been highlighted through the data sets Anscombe's quartet [@anscombe] and the Datasaurus Dozen [@datasaurpkg]. Each of the pairwise plots in these data sets have the same summary statistics but strikingly different information, as revealed when visualised. Visualisations can **tell** us what is important about our data and **how** it might diverge from pre-conceived notions or what we expect.

@@

However, uncertainty visualisation takes many forms today, It can be considered to expand the application of data visualisation towards making better decisions, and research is concerned about how the construction of plots and information contained might affect the resulting decisions. With the abundance of publications it is timely to consider a review of the state-of-the-art. In fact, there have already been several reviews published.

@@

# load libraries
library(strayr)
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)

>>>>

```{r}
#| echo: false
#| message: false
#| warning: false

# load Libraries
library(strayr)
library(tidyverse)
library(biscale)
library(RColorBrewer)
library(scales)
library(sf)

```
# Background

From entertainment choices to news articles to insurance plans, the modern citizen is so over run with information in every aspect of their life it can be overwhelming. In this overflow of information, tools that can effectively summarize information down into simple and clear ideas become more valuable. Information visualisations remain one of the most powerful tools for fast and reliable science communication.

There are many stages in our analysis that might benefit from the power of data visualisation, however this does not mean it is appropriately utilized or done with with effectiveness. Unlike an infographic, a statistical graphic is one of intent, and all enhancements made on a statistical graphic should be in pursuit of that goal (@Vanderplas2020). Visualization is an important step in exploratory data analysis and communication. At the exploration stage, it is utilized to **learn** what is important about a data set while the communication stage is utilized to **tell** others about what we have learned. Each of these visualisation goals utilize an important attribute of visualisation. In both these cases, visualisations are powerful tools that can help us understand **how** our data or results might diverge from pre-conceived notions or what we expect.

@@

```{r}
#| echo: false
#| message: false
# Generate data
ib_data <- tibble(x=rnorm(1000, 8, 4))

# Picture 1
p1 <- ib_data %>%
  summarise(avg = mean(x),
          conf_95a = quantile(x, probs=c(0.025)),
          conf_95b = quantile(x, probs=c(0.975))) %>%
  ggplot(aes(y="NA")) +
  geom_point(aes(x=avg)) +
  geom_errorbar(aes(xmin = conf_95a, xmax = conf_95b), width = 0.1) +
  scale_x_continuous(name = "INCHES OF SNOW",
                     breaks=seq(0,19),
                     labels= ggplot2:::interleave(as.character(c(seq(0,18, 2), 19)), rep("", 11))[c(0:19, 21)],
                     limits=c(0,19)) +
  theme_classic() +
  theme(axis.line.y=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        aspect.ratio=1/10)

# Picture 2
# Picture 3
# Picture 4
# Picture 5
# Picture 6
# Picture 7
# Picture 8
# Picture 9

```
However, uncertainty visualisation takes many forms today, It can be considered to expand the application of data visualisation towards making better decisions, and research is concerned about how the construction of plots and information contained might affect the resulting decisions. With the abundance of publications it is timely to consider a review of the state-of-the-art. In fact, there have already been several reviews published.

@@

### 2024-05-01 writingdumpster.qmd
small edit + figures for background

1. Introduction
- The development of graphs and charts has been linked to the development of coordinate systems and abstract representations of data
- the aim of most experimental research in statistical graphics focuses on the initial perception and graph comprehension, very little work has been done to understand the effect of charts and graphs on higher cognitive processes such as learning or analysis
1.1 Design of Statistical Graphs
- improved tech allow charts to be made easily for exploratory purposes
- consider purposes of presentation, entertainmen, analysis. Also considr a continum of utalitarian to artistry as a purpose.
- infographic tells a story (has chart junk) vs a statistical graphic is one of intent (and any visual enhancements should contribute to that aim)
- Points out that the ocean colour and state boundaries in the hurriance map only exist for geographic purposes. The map depicted here has the warnings information along with the cone.
1.2 Statistics Mapping using a Grammar of Graphics
- Systems that attempt to categorize charts based on their geometric representations generally make no effort to include all types of graphics, and they have difficulty accommodating charts that may fall into two or more categories. The classifications of graphics based on the underlying components and their relationships, as in the grammar of graphics developed byWilkinson, are more robust; they also provide an elegant framework for comparing different types of graphical representations separate from the underlying data structure. The former is like treating plots like creatures in a zoo, with a unique name for each, while the latter is analogous to having a phylogeny based on genetic data showing how plots are related.
- Has a visual explanation of the grammar of graphics.
- The grammar of graphics creates visual statistics (and conversely, allows plots to be considered statistics). The data plot can then be treated like other statistics: by imagining what the plot might look like in the absence of any structure, we can use the plot of the actual observed data to test for the likelihood of any perceived structure being significant.
- Using the grammar ofgraphics, it is easy for experimenters to compare different types of charts using the same data, as the underlying structure of the graph remains the same. 
- this approach to transformations and scales allows experimenters to easily test judgments made utilizing different axis transformations and color scales to compare perceptual accuracy

>>>>

1. Introduction
- The development of graphs and charts has been linked to the development of coordinate systems and abstract representations of data
- the aim of most experimental research in statistical graphics focuses on the initial perception and graph comprehension, very little work has been done to understand the effect of charts and graphs on higher cognitive processes such as learning or analysis
1.1 Design of Statistical Graphs
- improved tech allow charts to be made easily for exploratory purposes
- consider purposes of presentation, entertainmen, analysis. Also considr a continum of utalitarian to artistry as a purpose.
- infographic tells a story (has chart junk) vs a statistical graphic is one of intent (and any visual enhancements should contribute to that aim)
- Points out that the ocean colour and state boundaries in the hurriance map only exist for geographic purposes. The map depicted here has the warnings information along with the cone.
1.2 Statistics Mapping using a Grammar of Graphics
- Systems that attempt to categorize charts based on their geometric representations generally make no effort to include all types of graphics, and they have difficulty accommodating charts that may fall into two or more categories. The classifications of graphics based on the underlying components and their relationships, as in the grammar of graphics developed by Wilkinson, are more robust; they also provide an elegant framework for comparing different types of graphical representations separate from the underlying data structure. The former is like treating plots like creatures in a zoo, with a unique name for each, while the latter is analogous to having a phylogeny based on genetic data showing how plots are related.
- Has a visual explanation of the grammar of graphics.
- The grammar of graphics creates visual statistics (and conversely, allows plots to be considered statistics). The data plot can then be treated like other statistics: by imagining what the plot might look like in the absence of any structure, we can use the plot of the actual observed data to test for the likelihood of any perceived structure being significant.
- Using the grammar ofgraphics, it is easy for experimenters to compare different types of charts using the same data, as the underlying structure of the graph remains the same. 
- this approach to transformations and scales allows experimenters to easily test judgments made utilizing different axis transformations and color scales to compare perceptual accuracy
